Fast approximate Bayesian inference of HIV indicators using PCA adaptive Gauss-Hermite quadrature

### 2024-12-15 src/docs_paper-jtb/paper.Rmd
Fix to figure caption

The Naomi model was difficult to efficiently sample from using NUTS via `tmbstan`.
The lowest effective sample size was `r round(naomi_nuts$ess_min, 0)` (2.5% quantile `r round(naomi_nuts$ess_lower, 0)`, 50% quantile `r round(naomi_nuts$ess_median, 0)`, and 97.5% quantile `r round(naomi_nuts$ess_upper, 0)`; Panel \@ref(fig:worst-trace)A).
The largest potential scale reduction factor was `r round(naomi_nuts$rhat_max, 3)` (2.5% quantile `r round(naomi_nuts$rhat_lower, 3)`, 50% quantile `r round(naomi_nuts$rhat_median, 3)`, and 97.5% quantile `r round(naomi_nuts$rhat_upper, 3)`; Panel \@ref(fig:worst-trace)B).
Though inaccuracies remain possible, these diagnostics are sufficient to treat inferences obtained from NUTS as a gold-standard.

@@

(ref:total-variation) Under PCA, the proportion of total variation explained is given by the sum of the first $s$ eigenvalues over the sum of all eigenvalues. A typical rule-of-thumb is to include dimensions sufficient to explain 90% of total variation. In this case, for computational reasons, `r round(100 * tv_df$tv[8])`% was considered sufficient.

```{r total-variation, fig.cap="(ref:total-variation)"}
knitr::include_graphics("depends/total-variation.png")
```

@@

For the PCA-AGHQ quadrature grid, a Scree plot based on the spectral decomposition of $\hat {\mathbf{H}}_\texttt{LA}^{-1}$ (as defined in Equation \@ref(eq:hessian2)) was used to select the number of principal components to keep (Figure \ref{fig:total-variation}).
Keeping $s = 8$ principal components was sufficient to explain `r round(100 * tv_df$tv[8])`% of total variation.
The reduced rank approximation to the inverse curvature with this choice of $s$ was visually similar to the full rank matrix (Figure \@ref(fig:reduced-rank)).

The principal component (PC) loadings (Figure \{fig:pc-loadings}) provide interpretable information about which directions had the greatest variation.
Many of the first PC loadings are sums of two hyperparameters.
As such, there is some redundancy in the hyperparameter parameterisation, supporting the findings of Section \@ref(naomi-nuts) regarding correlation structure in the hyperparameter posterior.
It is exactly this correlation structure that PCA, and PCA-AGHQ, looks to utilise.

@@

Projecting the $3^8 = 6561$ PCA-AGHQ quadrature nodes onto each hyperparameter dimension, there was substantial variation in coverage by hyperparameter (Figure \ref{fig:nodes-samples-comparison}).
Approximately 12 hyperparameters had well covered marginals: greater than the 8 naively obtained with a dense grid, but nonetheless far fewer than the full 24.
Coverage was higher among hyperparameters on the logistic scale, and lower among hyperparameters on the logarithmic scale.
This discrepancy occurred due to logistic hyperparameters naturally having higher posterior marginal standard deviation than logarithmic hyperparameters (Figure \@ref(fig:marginal-sd)).

@@

Posterior inferences from GEB, GPCA-AGHQ and NUTS were compared using point estimates (Section \@ref(point-estimates)), distributional quantities (Section \@ref(distributional-quantities)), and exceedance probabilities (Section \@ref(exceedance))

>>>>

The Naomi model was difficult to efficiently sample from using NUTS via `tmbstan`.
The lowest effective sample size was `r round(naomi_nuts$ess_min, 0)` (2.5% quantile `r round(naomi_nuts$ess_lower, 0)`, 50% quantile `r round(naomi_nuts$ess_median, 0)`, and 97.5% quantile `r round(naomi_nuts$ess_upper, 0)`; Panel XA).
The largest potential scale reduction factor was `r round(naomi_nuts$rhat_max, 3)` (2.5% quantile `r round(naomi_nuts$rhat_lower, 3)`, 50% quantile `r round(naomi_nuts$rhat_median, 3)`, and 97.5% quantile `r round(naomi_nuts$rhat_upper, 3)`; Panel XB).
Though inaccuracies remain possible, these diagnostics are sufficient to treat inferences obtained from NUTS as a gold-standard.

@@

```{r total-variation, fig.cap=paste0("Under PCA, the proportion of total variation explained is given by the sum of the first $s$ eigenvalues over the sum of all eigenvalues. A typical rule-of-thumb is to include dimensions sufficient to explain 90% of total variation. In this case, for computational reasons, ", round(100 * tv_df$tv[8]), "% was considered sufficient.")}
knitr::include_graphics("depends/total-variation.png")
```

@@

@@

For the PCA-AGHQ quadrature grid, a Scree plot based on the spectral decomposition of $\hat {\mathbf{H}}_\texttt{LA}^{-1}$ was used to select the number of principal components to keep (Figure \ref{fig:total-variation}).
Keeping $s = 8$ principal components was sufficient to explain `r round(100 * tv_df$tv[8])`% of total variation.
The reduced rank approximation to the inverse curvature with this choice of $s$ was visually similar to the full rank matrix (Figure \ref{fig:reduced-rank}).

The principal component (PC) loadings (Figure Appendix) provide interpretable information about which directions had the greatest variation.
Many of the first PC loadings are sums of two hyperparameters.
As such, there is some redundancy in the hyperparameter parameterisation, supporting the findings of the Appendix regarding correlation structure in the hyperparameter posterior.
It is exactly this correlation structure that PCA, and PCA-AGHQ, looks to utilise.

Projecting the $3^8 = 6561$ PCA-AGHQ quadrature nodes onto each hyperparameter dimension, there was substantial variation in coverage by hyperparameter (Figure \ref{fig:nodes-samples-comparison}).
Approximately 12 hyperparameters had well covered marginals: greater than the 8 naively obtained with a dense grid, but nonetheless far fewer than the full 24.
Coverage was higher among hyperparameters on the logistic scale, and lower among hyperparameters on the logarithmic scale.
This discrepancy occurred due to logistic hyperparameters naturally having higher posterior marginal standard deviation than logarithmic hyperparameters (Figure Appendix).

@@

Posterior inferences from GEB, GPCA-AGHQ and NUTS were compared using point estimates (Section \ref{sec:point-estimates}), distributional quantities (Section \ref{sec:distributional-quantities}), and exceedance probabilities (Section \ref{sec:exceedance}).
